%% ========================================================================
%%							Acknowledgement
%% ========================================================================


\chapter*{Acknowledgement}
\label{cha:Acknowledgement}
Many people accompanied me on my way to the graduation of this study, whose acquaintance I enjoyed very much. 
I would also like to thank my supervisor Prof. Thorsten Rheinländer very much for his patience which he has given me in the course of the longer supervision time and for his outstanding support with each of my questions.
My colleagues at work also deserve a big thank you for always having an open ear for all my questions. I would especially like to thank my boss Dietmar Hareter for his continuous expertise and mental support, which has certainly been an essential factor in helping me to complete all my open tasks. 
I particularly want to thank my fellow students Birgit Bauer and Markus Kakac, who have made a significant and essential contribution to the fact that I always look back on my studies with pleasure. 
An important constant in my life that supports me in all my decisions and is always there for me is my beloved girlfriend Claudia Hilsberg who helped me never lose sight of the goal.
All that I have achieved so far would not have been possible without the great support of my parents. I look back with great joy on the talks we had together about all those things that we could do in the future, but unfortunately it was not meant to be that we could celebrate my graduation together. You will always be an important part of my life. 

